Wide--Field Effects in Redshifted 21~cm Power Spectra

### 2015-03-12 ms.tex
Add preliminary section 1 "Introduction"
From 85866d7:
============

In ms.tex,

(*) Add preliminary section 1 "Introduction"
(*) Make minor changes to "Rseults" section
(*) Make minor changes to "MWA Observations" section

@@

Typically, an antenna is most sensitive towards its primary field of view relative to the rest of the sky, which in conjunction with the aforementioned wide--field effects results in a characteristic ``pitchfork'' signature in the delay spectrum.

@@

The MWA instrument configuration, EoR observations, and data analysis used in this study are already described in \citet{thy15}. In order to reduce thermal fluctuations while maintaining coherence, it is essential to average independent data sets obtained over the same region of sky with identical beamformer settings. Hence, we select a subset of MWA snapshots each of duration 112~seconds obtained over different nights which are aligned to within 72~seconds of each other in {\it local sidereal time} (LST) around a mean LST of 0.04~hours with the MWA tile beam pointed at zenith. The database contains 14 snapshots satisfying these criteria. Two of these snapshots were found to contain amplitude and phase artifacts for a significant duration across different baselines. Hence, they have been excluded from our analysis.

@@

We focus on the level of foreground power near the horizon limits in the data. Typically, the power near the negative horizon limit is seen with a signal--noise ratio (SNR) $\sim$~10--100, while that around the positive horizon limit is $\sim$~100--1000. The models of \citet{thy15} have noted that the foreground power near the horizon limits is due to the nature of wide--field measurements, and is predominantly composed of diffuse emission especially on baseline lengths $\lesssim 100$~m. Based on the morphological agreement between the data and the models, we conclude the features noted in our current analysis are a robust detection of the {\it pitchfork} signature predicted in \citet{thy15}.

>>>>

The epoch of reionization (EoR) commenced following the formation of the first stars and galaxies. It is characterized by a period of non--linear growth of matter density perturbations and astrophysical evolution in the Universe's history. Detection of redshifted 21~cm radiation of H{\sc i} from this epoch is one of the most promising probes of the non--linear evolution of large scale structure characteristic of this epoch \citep{sun72,sco90,mad97,toz00,ili02}.

While sensitive instruments such as the Square Kilometre Array (SKA) with the capability of direct imaging of redshifted H{\sc i} become operational, numerous pathfinders and precursors such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, and the planned Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) have enough sensitivity for a statistical detection of the signal \citep{bow06,par12a,bea13,dil13,thy13,pob14} and most of these are already operational.

The primary challenge to detection of cosmological redshifted H{\sc i} from the EoR comes from continuum emission from Galactic and extragalactic foreground objects, relative to which the desired signal is $\sim 10^4$ times weaker \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. Since the foregrounds and the signal have inherent differences in spatial isotropy and spectral smoothness \citep{mor04,mor06,bow09,liu11,par12b,dil13,pob13}, a detailed characterization of foreground emission has become essential \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,dil14,liu14a,liu14b,thy13,thy15}.

A recent study by \citet{thy15} using instrument and all--sky foreground models has shed new light on the significant role played by the interplay between foreground emission and the nature of wide--field measurements, typical of most modern EoR experiments. This paper presents a follow--up analysis of that study. By using a larger volume of MWA data obtained independently over different nights pointed at the same region of sky, we confirm with high significance the presence of certain key characteristics (referred to as the {\it pitchfork} signature) of wide--field measurements predicted in the preceding study.

\S\ref{sec:wide-field} gives an overview of the role of wide--field measurements in the delay spectral domain and the predicted {\it pitchfork} signature. In \S\ref{sec:MWA}, we describe the data analysis procedure used to improve the dynamic range of delay spectra of MWA data. We describe the results and confirm the presence of the predicted wide--field effects in \S\ref{sec:results}. \S\ref{sec:summary} summarizes our findings.

@@

Typically, an antenna is most sensitive towards its primary field of view relative to the rest of the sky, which in conjunction with the aforementioned wide--field effects results in a characteristic {\it pitchfork} signature in the delay spectrum.

@@

The MWA instrument configuration, EoR observations, and data analysis used in this study are already described in \citet{thy15} and references therein. In order to reduce thermal fluctuations while maintaining coherence, it is essential to average independent data sets obtained over the same region of sky with identical beamformer settings. Hence, we select a subset of MWA snapshots each of duration 112~seconds obtained over different nights which are aligned to within 72~seconds of each other in {\it local sidereal time} (LST) around a mean LST of 0.04~hours with the MWA tile beam pointed at zenith. The database contains 14 snapshots satisfying these criteria. Two of these snapshots were found to contain amplitude and phase artifacts for a significant duration across different baselines. Hence, they have been excluded from our analysis.

@@

We focus on the peaks near the horizon limits in the data. Typically, the power near the negative horizon limit is seen with a signal--noise ratio (SNR) $\sim$~10--100, while that around the positive horizon limit is $\sim$~100--1000. The models of \citet{thy15} have noted that the foreground power near the horizon limits is due to the nature of wide--field measurements, and is predominantly composed of diffuse emission especially on baseline lengths $\lesssim 100$~m. Based on the morphological agreement between the data and the models, we conclude the features noted in our current analysis are a robust detection of the {\it pitchfork} signature predicted in \citet{thy15}.
